A flexible framework to prototype, develop and manage ubiquitous applications for intelligent environments

### 2013-10-03 trunk/ labtempo-ubicomp-projects/UbicompMiddleware/doc/PerCom2014/percom2014_conf.tex
New image
Started smartandroid location subsection

    %[talk about stub intantiation calls]
    %TODO: example of CV and OP? Maybe it is worth add some on the text
    To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
    \begin{enumerate}
     \item The UbiApp calls a method from the stub 
     \item The call is forwarded to the RA instance through the network in the second step
     \item This instance can interact with the device (the lamp, in case)
     \item The result or the acknowledge is then returned to the caller
    \end{enumerate}

    The second stage happens when the first RA is already subscribed to the second and a notification routine is triggered, what was represented by the Figure~\ref{fig:pubsub}.b and works as follows:
    \begin{enumerate}
     \item The device has its context changed, what may be caused directly (e.g., a user turning the lamp on) or through the RA (i.e. an application, another RA, or rule, calling some RA's OP which caused the change of context)
     \item Then the RA that represents the device notifies all subscriptors to that CV (e.g., ``Is on'' from this Lamp RA)
     \item The subscriptor may call some callback method of the UbiApp
    \end{enumerate}

@@

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The SmartAndroid middleware contemplates the possility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a SmartAndroid Market~\cite{google2013}, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do.
% TODO: talk about market place
[security concern on store: helps the discovery of malware before the user download]

  \subsection{Distributed Components Model}
%   [implementation of RA and services]

  [initialization]
%   O SmartAndroid é um aplicativo que inicializa o SGAR, repositório e serviços de comunicação. Este aplicativo deve ser executado em uma única unidade do ambiente, deixando o Deamon de Comunicação em estado de espera até receber requisições de outras aplicações. Em outras unidades (clientes) constam os aplicativos instalados com a biblioteca do SmartAndroid, a qual é necessária para executar o serviço de comunicação e \textit{proxys} dos serviços do SGAR. Assim cada unidade pode consultar sobre IARs, através do SDR, obtendo seus ODARs, cada um com o respectivo RANS, o qual é utilizado pelo \textit{proxy} como referência de acesso.

  with the RANS 
  RPC search

  [annotation]

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example Applications} \label{sec:evaluation}
Some applications have been built to explore the potential of the framework for building UbiApps. Although the example applications are simple, and alone does not justify the deploy of the SmartAndroid in the environment, they have allowed us to validate the features provided by the framework and achieved by the SmartAndroid.

% Among these applications we will

% the SmartLiC, the MediaFollowMe, remote control type applications where resources are controlled remotely by another resource (in this case, a smartphone ), among other applications. A future work is to install a system in Ambi home care health, as in ~ \ cite {} carvalho2010sciads.

@@

  \subsection{SmartLiC}
  % [smartlic]
  The Smart Light Control (SmartLic) application aims at the clever use of lighting in homes in order to save power consumption, since it is responsible for the largest range of consumption in residential environments. The application uses presence sensors to identify the location of people inside each room, and decide which lamps should be turned on and off. Hence it prevents the forgetting of lamps turned on, and may lit some lamp if a user enters a room at night and if nobody is at sleep in that room. 
  % ref to the 'largest range of consumption in residential environments'?

  The [IPGAP] has allowed us to test this applications. We instantiated lamps simulators for each room and an avatar for a user, and also a simulator for day light sensor. Next context rules were created by filling a skeleton of a rule written in json, what caused the instantiation of a CI and actuator for each room.  If the avatar is dragged, in the [IPGAP], from a room and dropped in another, it ma
%   O Código 2 ilustra uma possível implementação da interpretação de contexto pela aplicação \mbox{SmartLiC}. Na etapa de inicialização, a aplicação utiliza o SLR para localizar os cômodos (Linha 2), e obter a lâmpada e o sensor de presença (``lampAR'' e ``presAR'') de cada um, utilizando a busca do SDR (Linhas 4 e 5). Para evitar a repetição do código da regra, suas referências e a respectiva ação a ser desempenhada (desligar a lâmpada referenciada -- ``lampAR.turnOff'') são adicionadas na lista de configurações (``configList'') (Linha 6). A regra é criada em ``rule'' como: ``\mbox{\emph{lâmpada ligada E o cômodo desocupado POR T$_{max}$ unidades de tempo}}'', e desempenha uma relação lógica entre as diferentes VCs e a temporização da regra (Linha 7). Em seguida, subscreve-se às VCs dos ARs de interesse, começando o processamento da regra (Linha 8).

@@

  \subsection{Remote Controls}
  Considering that the users would want to have some control over the resources within the IE, we have implemented the remote controls to interact with them specifically.

  \begin{figure}[!t]
  \centering
  \includegraphics[width=3.5in]{figures/beagle-lamp.pdf}
  \caption{Remote Control} %[IPGAP]
  \label{fig:remotecontrol}
  \end{figure}

>>>>

@@

%proposta de uso de marketplace
% A exposição de dados por meio de \textit{Content Provider} permite a integração de aplicativos externos ao servidor. Estes aplicativos podem ser dispostos em um \textit{marketplace} (http://developer.android.com/distribute/open.html) permitindo que tanto usuários comuns quanto programadores incrementem o ambiente com novas aplicações ubíquas. 
The SmartAndroid middleware contemplates the possility of running multiple applications in the same IE. Thus, focusing on the user and inspired by the Google Play -- which is a application market for Android devices --, we have as a future work the creation of a SmartAndroid Market~\cite{google2013}, where users will be able to easily download and start different applications. Also similar to the Google Play, a Manifest file will require permission for accessing the domains, what will allow users to maintain control over what off-the-shelf applications have permission to know and to do. Within the market, the UbiApp could be tested towards discovering any possible harm that the application could do and any possible inefficiency that could consume too much the energy resources of the host device.
% TODO: talk about market place
% [security concern on store: helps the discovery of malware before the user download]

  \subsection{Location}

  [talk about location]

  The devices' location may be acquired through many different approaches as radio frequency signals such as RFID tags \cite{?}, GSM, Bluetooth and Wifi or even a smart floor which can locate people and objects by a pressure-sensing floor. However, the SmartAndroid doesn't implement a specific technique to locate resources, in order to keep the framework generic. The SmartAndroid API deal with location in a hi-level way making the data acquirement transparent to it. This aspect is important to keep the system flexible to any approach, allowing the developer to use different implementations.

  \subsection{Communication Support}
  %[talk about stub intantiation calls]
  %TODO: example of CV and OP? Maybe it is worth add some on the text
  To perform any interaction with a RA that is not locally instantiated, the UbiApp must use a stub (also called proxy) that is bound to the referred RA. This stub implements the same interface as the RA, i.e. the same CVs and OPs, which are methods, from a Object-Oriented Programming perspective. With the stub on hands, the application can invoke remote methods, as if they were locally invoked. These remote methods may be to retrieve a CV or to trigger an OP. The Figure~\ref{fig:rpc} represents this direct, or synchronous, communication. The steps to achieve a direct communication, as represented by the figure, are as follows:
  \begin{enumerate}
    \item The UbiApp calls a method from the stub 
    \item The call is forwarded to the RA instance through the network in the second step
    \item This instance can interact with the device (the lamp, in case)
    \item The result or the acknowledge is then returned to the caller
  \end{enumerate}

  The second stage happens when the first RA is already subscribed to the second and a notification routine is triggered, what was represented by the Figure~\ref{fig:pubsub}.b and works as follows:
  \begin{enumerate}
    \item The device has its context changed, what may be caused directly (e.g., a user turning the lamp on) or through the RA (i.e. an application, another RA, or rule, calling some RA's OP which caused the change of context)
    \item Then the RA that represents the device notifies all subscribers to that CV (e.g., ``Is on'' from this Lamp RA)
    \item The subscriber may call some callback method of the UbiApp
  \end{enumerate}

% 
%   \subsection{Distributed Components Model}
% %   [implementation of RA and services]
%   
%   [initialization]
% %   O SmartAndroid é um aplicativo que inicializa o SGAR, repositório e serviços de comunicação. Este aplicativo deve ser executado em uma única unidade do ambiente, deixando o Deamon de Comunicação em estado de espera até receber requisições de outras aplicações. Em outras unidades (clientes) constam os aplicativos instalados com a biblioteca do SmartAndroid, a qual é necessária para executar o serviço de comunicação e \textit{proxys} dos serviços do SGAR. Assim cada unidade pode consultar sobre IARs, através do SDR, obtendo seus ODARs, cada um com o respectivo RANS, o qual é utilizado pelo \textit{proxy} como referência de acesso. 
% 
% 
%   with the RANS 
%   RPC search 
% 
%   
%   [annotation]

%   \subsection{Resource Agent [?????]}

%   Inside the RA,

  We have provided a simple implementation for the pub-sub mechanism, which was effective to evaluate our approach. The Figure~\ref{fig:ra-intercept} shows how we have implemented this mechanism using the RPC. In the example, two RAs (TV and Alarm Clock) subscribe to the Stove RA (step 1). Every RA keeps track of its stakeholders, where the stakeholder list keeps the RANS and the CV of interest. When a remote control (or another actuator) calls a OP that changes the state of the Stove RA (step 2), this is intercepted (step 3) and causes the selection of the stakeholders (step 4) followed by the notification (step 5) that also uses the RPC.

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.4in]{figures/publish-subscribe-ar.pdf}
  \caption{Resource Agent} 
  \label{fig:ra-intercept}
  \end{figure}

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Example Applications} \label{sec:evaluation}
Some applications have been built to explore the potential of the framework for building UbiApps. Although the example applications are simple, and alone does not justify the deploy of the SmartAndroid in the environment, they have allowed us to validate the features provided by the framework and achieved by the SmartAndroid. It also demonstrated the possibility of concomitant running of UbiApps, what similar to an operation system where different processes run separately but in the same system. Withing the SmartAndroid, UbiApps run separately but in the IE.% and can share resources,

% In this section we will elaborate over the SmartLiC application, and next we will briefly talk about the others.
In the subsections that follows we will elaborate over some UbiApps that have been developed, or are still in development, in the course of the project.
% the SmartLiC, the MediaFollowMe, remote control type applications where resources are controlled remotely by another resource (in this case, a smartphone ), among other applications. A future work is to install a system in Ambi home care health, as in ~ \ cite {} carvalho2010sciads.

@@

  \subsection{SmartLiC}
%   primeiramente vc cria um skeleton para a regra
% 
% depois vc pega cada cômodo e de cada um deles faz
  % pega a lampada 
%   pega o sensor de presença
%   cria um CI com a regra e as devidas referencias
%   instancia atuador desligador de lampadas
%   subscreve atuador no ic

  % [smartlic]
  The Smart Light Control (SmartLic) application aims at the clever use of lighting in homes in order to save power consumption, since it is responsible for the largest range of consumption in residential environments. The application uses presence sensors to identify the location of people inside each room, and decide which lamps should be turned on and off. Hence it prevents the forgetting of lamps turned on, and may lit some lamp if a user enters a room at night and if nobody is at sleep in that room. 
  % ref to the 'largest range of consumption in residential environments'?

  The [IPGAP] has allowed us to test this application. We instantiated lamp simulators for each room and an avatar for a user, and also a simulator for day light sensor. Next context rules were created for each room of the house. %Next context rules were created by filling a skeleton of a rule written in json, what caused the instantiation of a CI and actuator for each room.  If the avatar is dragged, in the [IPGAP], from a room and dropped in another, and the specified time passes, the the lamp is turned off.
  \begin{enumerate}
   \item Create a context rule skeleton in JSON
   \item Get the rooms from the house
   \item Get the lamp and presence sensor of each room and fill the rule expression skeleton
   \item Instantiate a CI with the rule expression
   \item Instantiate a actuator to perform the actions declared in the context rule skeleton
   \item Bind the CI and the actuator
   \item Repeat this for each room
  \end{enumerate}

  The event of a user entering some room is triggered when the avatar is dragged from one room and dropped on another. The CI has a timer associated with the expression so, when the rule is valid, the timer is fired. If the context change and the rule is invalidated the timer is stopped, otherwise the evaluation is completed with a valid rule and the subscribers must be notified. Then the actuator turns off the light of the first room.

  There were also rules composed for turning on the light of the room that the user enters if it is day (based on a day light sensor simulator) and if nobody is at sleep in the room (based on another CI).

  The SmartLiC allowed us to validate two features of the RLS, which are the discovery of the rooms and the RAs of some specific type are within the room (i.e. the lamps and presence sensors). In this UbiApp we also have created a set of context rules with timers using the context rules parser (including CI and actuators). And the [IPGAP] were well explored in this implementation, since we has shown the triggering of many events in parallel and multiple users cohabiting the same IE.
% 
%   After the start of the context rules, 
%   \begin{enumerate}
%    \item On event received, get the sender, the 
%   \end{enumerate}

%   O Código 2 ilustra uma possível implementação da interpretação de contexto pela aplicação \mbox{SmartLiC}. Na etapa de inicialização, a aplicação utiliza o SLR para localizar os cômodos (Linha 2), e obter a lâmpada e o sensor de presença (``lampAR'' e ``presAR'') de cada um, utilizando a busca do SDR (Linhas 4 e 5). Para evitar a repetição do código da regra, suas referências e a respectiva ação a ser desempenhada (desligar a lâmpada referenciada -- ``lampAR.turnOff'') são adicionadas na lista de configurações (``configList'') (Linha 6). A regra é criada em ``rule'' como: ``\mbox{\emph{lâmpada ligada E o cômodo desocupado POR T$_{max}$ unidades de tempo}}'', e desempenha uma relação lógica entre as diferentes VCs e a temporização da regra (Linha 7). Em seguida, subscreve-se às VCs dos ARs de interesse, começando o processamento da regra (Linha 8).

@@

  \subsection{Remote Controls}
  Considering that the users would want to have some control over the resources within the IE, we have implemented the remote controls to interact with them specifically. The same way that users nowadays have a TV control that enable them to have full control over the device with little effort, using the framework that task will be easily extended to the diverse range of devices present in the IE.

  For implementing a lamp remote control, as shown on the Figure~\ref{fig:remotecontrol}, we are using a prototyping board called BeagleBone~\cite{Beagle2012}. The board, on one side, is capable of turning a lamp on and off and, on the other side, it implements a RA for the lamp which is registered in the IE. A graphical interface of a lamp can remotely call these OPs of the lamp, and it is also subscribed to the Lamp RA, what allows the remote control being aware of the lamp status.

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/beagle-lamp.pdf}
  \caption{Remote Control} %[IPGAP]
  \label{fig:remotecontrol}
  \end{figure}
